---
title: "Homework 2"
author: "Diego Saráchaga"
date: "February 11, 2019"
output:
  html_document: default
  pdf_document: default
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(magrittr)
library(lubridate)

tv <- read_csv('https://github.com/rfordatascience/tidytuesday/raw/master/data/2019/2019-01-08/IMDb_Economist_tv_ratings.csv')
```

### Exercise 2

----

#### Part a)
__R Code:__   
```{r prelim, error = TRUE, message = FALSE, warning = FALSE, eval = TRUE}
tv %>% filter(genres == "Action,Adventure,Drama", date > ymd("17/1/1")) %>% 
  group_by(title, seasonNumber, date) %>% 
  arrange(desc(av_rating))
```

__Conclusion:__   
After 2017 in the category _"Action,Adventure,Drama"_, the show that had the highest average rating was _Game of Thrones_ in its __7th season__, with an __average rating of 9.3__.

   
----

#### Part b)
__R Code:__   
```{r prelim2, error = TRUE, message = FALSE, warning = FALSE, eval = TRUE}
dat <- filter(tv, title == "Law & Order" | title == "Law & Order: Special Victims Unit")
ggplot(dat, aes(date, av_rating, color = title)) + 
  geom_point() + geom_line() + geom_smooth() +
  ggtitle(expression("Date and Avg Rating"~italic("Law & Order")~"vs"~italic("Law & Order: Special Victims Unit"))) + 
  xlab("Date") + 
  ylab("Avg Rating")
```
---
__Comments:__   
First, we can see that _Law & Order_ seasons were from __1991 to 2010__, while _Law & Order: Special Victims Unit_ started on __2000 and continues till 2018__.  
Then, in the plot we can see that from its begining _Law & Order: Special Victims Unit_ has had a __higher rating__ than _Law & Order_ when they where both on air. However, on __season 16 (year 2015)__, _Law & Order: Special Victims Unit_ had its __worst rating__, that is also the worst rating if we consider the history of both shows.

----

#### Part c)
__R Code:__   
```{r prelim3, error = TRUE, message = FALSE, warning = FALSE, eval = TRUE}
dat <- filter(tv, title == "Criminal Minds")
ggplot(dat, aes(date, av_rating, color = title)) + 
  geom_point() + geom_line() + geom_smooth() +
  ggtitle(expression("Date and Avg Rating"~italic("Criminal Minds")))+
  xlab("Date") + 
  ylab("Avg Rating")
```

---
__Comments:__   
First, we can see that _Criminal Minds_ seasons were from __2006 till 2018__.  
Then, in the plot we can see that from the begining until the year 2016 the average rating __was mostly declining__, with the exception of season 2. Moreover, in 2016 _Criminal Minds_ reached its worst average rating in the show's history.   
However, the last 3 ratings __have increased__ compared to the previous ratings.   
Finally, we can see that in 2018 (last 2 ratings), __the show presented its highest historically average rating__.
